Proposed Solution

### 2016-11-11 Ennan-Proposal/motivation.tex
stuff

We believe there are two main obstacles 
why we cannot simply apply the existing automatic 
tools and techniques to verification of configuration files are:
\begin{itemize}
\item a lack
of a specification which would describe properties of configuration files
\item a program structure of configuration files -- they
are mainly a sequence of entries assigning some value to system variables. 
\end{itemize}
The language in which configuration files are written does 
not adhere to a specific grammar or syntax. In particular, the
entries in configuration files are untyped. Moreover, there are surprisingly few rules specifying constraints on entries and there
is no explicit structure policy for the entries.
We believe that those are the reasons why there is no automated verification of configuration 
files although the system community expressed that that would be highly
desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

@@

The goal of this proposal is to develop a fully {\bf {automated 
verification  framework for general software configurations}}. We plan to
overcome the above obstacles by first automatically inferring a
specification for configuration files. It is unrealistic to expect the 
users to write an entire specifications for configuration files on their own. 
This process can easily lead to incomplete or even contradictory 
specifications. Instead, we will learn specification from 
a large sample of 
configuration files. A learning process will take as input this 
sample. The process will be language-agnostic and should work for any kind 
of configuration files,  but all of the files in the sample need to be of 
the same kind (such as MySQL or 
HTTPD configuration files). From that sample, we will learn an abundant set of rules 
specifying various properties that hold on the given sample. The rules, in general, 
specify which properties variables in configuration files need to satisfy. One can see this learning process as 
a way of deriving  a specification for configuration files. With these  rules we can efficiently check 
the correctness of the configuration files of interest and detect 
potential errors. Errors are reported if the configuration file does not 
adhere to a specification.

@@

Building such an automatic verification framework for
configuration files requires addressing several challenges. 
First, in the process of inferring a specification, we have to 
analyze mainly with an untyped, unstructured sequence of assignments.
Thus, we need to develop a suitable language model. We do that by 
``guessing'' a type of variables and then deriving formulas that
describe relationships between these variables. However, the
type of a variable cannot always be fully determined 
from a single value. 
An entry {\tt general\_log = 1} in a MySQL configuration file assigns an integer to the 
variable {\tt general\_log}. However, 1 here is a Boolean variable,
which denotes that there will be a general log file.
Another example, which often appears in practice, shows how variable names and 
their ``types'' can easily lead to an error. Consider the
following entry {\tt general\_log = /var/log/mysql/mysql.log}. Although this statement 
alone might look correct, this is actually a typing error and as a result MySQL log cannot be correctly written.

Second, we will learn patterns describing variables and their relations. We will associate with each type a set of very general templates. The user does not need to provide any templates -- they are internally associated to the types.
Nevertheless, in addition to those general templates, we still need specific algorithms to learn
rules that cannot be easily templated (such as ordering rules).

>>>>

We believe there are two main obstacles 
why we cannot simply apply the existing automatic 
tools and techniques to verification of configuration files are:
\begin{itemize}
\item a lack
of a specification which would describe properties of configuration files
\item a program structure of configuration files -- they
are mainly a sequence of entries assigning some value to system keywords. 
\end{itemize}
The language in which configuration files are written does 
not adhere to a specific grammar or syntax. In particular, the
entries in configuration files are untyped. Moreover, there are surprisingly few rules specifying constraints on entries and there
is no explicit structure policy for the entries.
We believe that those are the reasons why there is no automated verification of configuration 
files although the system community expressed that that would be highly
desirable~\cite{wang04automatic, zhang14encore, xu15systems}.

@@

The goal of this proposal is to develop a fully {\bf {automated 
verification  framework for general software configurations}}. We plan to
overcome the above obstacles by first automatically inferring a
specification for configuration files. It is unrealistic to expect the 
users to write an entire specifications for configuration files on their own. 
This process can easily lead to incomplete or even contradictory 
specifications. Instead, we will learn specification from 
a large sample of 
configuration files. A learning process will take as input this 
sample. The process will be language-agnostic and should work for any kind 
of configuration files,  but all of the files in the sample need to be of 
the same kind (such as MySQL or 
HTTPD configuration files). From that sample, we will learn an abundant set of rules 
specifying various properties that hold on the given sample. The rules, in general, 
specify which properties keywords in configuration files need to satisfy. One can see this learning process as 
a way of deriving  a specification for configuration files. With these  rules we can efficiently check 
the correctness of the configuration files of interest and detect 
potential errors. Errors are reported if the configuration file does not 
adhere to a specification.

@@

Building such an automatic verification framework for
configuration files requires addressing several challenges. 
First, in the process of inferring a specification, we have to 
analyze mainly with an untyped, unstructured sequence of assignments.
Thus, we need to develop a suitable language model. We do that by 
``guessing'' a type of keywords and then deriving formulas that
describe relationships between these keywords. However, the
type of a keyword cannot always be fully determined 
from a single value. 
An entry {\tt general\_log = 1} in a MySQL configuration file assigns an integer to the 
keyword {\tt general\_log}. However, 1 here is a Boolean variable,
which denotes that there will be a general log file.
Another example, which often appears in practice, shows how keyword names and 
their ``types'' can easily lead to an error. Consider the
following entry {\tt general\_log = /var/log/mysql/mysql.log}. Although this statement 
alone might look correct, this is actually a typing error and as a result MySQL log cannot be correctly written.

Second, we will learn patterns describing keywords and their relations. We will associate with each type a set of very general templates. The user does not need to provide any templates -- they are internally associated to the types.
Nevertheless, in addition to those general templates, we still need specific algorithms to learn
rules that cannot be easily templated (such as ordering rules).

### 2016-11-11 Ennan-Proposal/prelim.tex
stuff

From ConfigC we drew the main inspiration for this proposal. We now briefly
outline how does ConfigC work. It first automatically
analyzes a dataset of correct configuration files to derive
 rules for describing a specification. Before that process even starts,
we identified several basic types, 
such Boolean, Integer, File. With each of these types we associated a 
list of templates. We then analyze the given training data set of 
correct configuration files and derive rules describing variable 
relations. We illustrate how it is done on two fragments of configuration 
files given in Fig.~\ref{fig:twoFiles}.  When analyzing
a file, we first do type-checking, \ie we assign a type to every variable. After that we instantiate each template with the typed variables and we save the learned set 
of rules. For instance, with the integer type we only associated comparison templates ($X < Y, X = Y, \ldots$). After 
analyzing File\_1, the set of learned rules will contain the following rules:
\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
max_connections=300
mysql.max_persistent=200
max_connections > mysql.max_persistent
\end{lstlisting}
The set of learned rules is changing with every file that we analyze: contradicting 
files are removed and new rules are added. In practice, this process is done 
in parallel.

>>>>

From ConfigC we drew the main inspiration for this proposal. We now briefly
outline how does ConfigC work. It first automatically
analyzes a dataset of correct configuration files to derive
 rules for describing a specification. Before that process even starts,
we identified several basic types, 
such Boolean, Integer, File. With each of these types we associated a 
list of templates. We then analyze the given training data set of 
correct configuration files and derive rules describing variable 
relations. We illustrate how is this done on two fragments of 
configuration files given in Fig.~\ref{fig:twoFiles}.  When analyzing
a file, we first assign a type to every 
variable. After that we instantiate each template with the typed 
variables and we save the learned set 
of rules. For instance, with the integer type we only associated comparison templates ($X < Y, X = Y, \ldots$). After 
analyzing File\_1, the set of learned rules will contain the following rules:
\begin{lstlisting}[language=C, xleftmargin=.01\textwidth]
max_connections=300
mysql.max_persistent=200
max_connections > mysql.max_persistent
\end{lstlisting}
The set of learned rules is changing with every file that we analyze: contradicting 
files are removed and new rules are added. In practice, this process is done 
in parallel.
